---
title: "Egg composite CVD study"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "tableone", "survival", "mice")
sapply(pacs, require, character.only = TRUE)

# Read the list of imputed data -------------------------------------------

imputed_data_list <- readRDS("./Data/imputed_data_list.rds")

# First imputed data
mi1 <- imputed_data_list[[1]]

```

## Aim

* Examine associations between egg and meat intake and incident cardiovascular disease (CVD)
* Examine whether egg intake modifies the association between meat intake and CVD
  * i.e., whether egg and meat intake interact in their effect on CVD incidence

## Outcome

* Any of the following incident CVD cases:
  * Ischemic heart disease
  * Stroke/transient ischemic attack
  * Acute myocardial infarction
  * Congestive heart failure (in 27 chronic conditions) or heart failure and non-ischemic heart disease (in 30 chronic conditions)
  * Atrial fibrillation
* For disease definition, see [Chronic Conditions Data Warehouse](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
  * Note that the definitions may differ between 27 chronic conditions (available in 2008-2020 data) and 30 chronic conditions (2021-2022)

## Exposure of interest

* Dietary intake, measured via the AHS baseline questionnaire

* Egg intake frequency
  * Categorized into four groups based on frequency of intake

* Meat intake, gram/day
  * Energy-adjusted
  
* Multiplicative interaction terms between egg intake category and meat intake (g/d) will be assessed in Cox proportional hazards models

## Datasets

* Medicare data
  * For details regarding Medicare data, see [AHS-2 Medicare Linkage](https://github.com/keijioda/ahs_medicare_linkage/blob/main/summary.md) repository.
  
  * Master Beneficiary Summary File (MBSF), 2008-2022
    * Contains beneficiary characteristics and enrollment information
    
  * Chronic Conditions file (CC), 2008-2022
    * Contains the first occurrence date of [27 (or 30) specific chronic conditions](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
    * Used to identify prevalent/incident cases of CVD and
    * to identify comorbidities, based on ICD-9 and ICD-10 codes
    
  * Both files include n = 46,897 unique subjects across years, after excluding 
    * Gender/DOB mismatch with AHS-2 data
    * Dupulicate beneficiary IDs and SSNs

* AHS-2 baseline data: n = 96,144
  
* After merging Medicare and AHS-2 data, there were n = 43,467 subjects
  * Participants who opted out from the study or those who live outside the U.S were already excluded at this point

## Inclusion/exclusion criteria

* Medicare beneficiaries who did not reach the age of 65 between 2008 and 2022 (e.g., younger beneficiaries with disabilities or end-stage renal disease) were excluded (n = 1238), resulting n = 42,229.

* n = 106 subjects with extreme BMI (<16 or >60), according to AHS questionnaire, were excluded, resulting n = 42,123.

* Unverified dates of deaths
  * Medicare data include a variable (`VALID_DEATH_DT_SW`) indicating whether a beneficiary's day of death has been verified by the Social Security Administration or the Railroad Retirement Board.
  * There were 23 unverified death dates. Excluding these resulted in n = 42,100.

* Prevalent CVD cases
  * At this point of analysis, prevalent cases were defined as those who were diagnosed with any of the five CVD conditions prior to their AHS-2 enrollment
  * There were 3,542 such prevalent cases. Excluding these prevalence resulted in n = 38,558
  
## Incident CVD cases

* Among n = 38,558 subjects, there were 12,890 incident CVD cases (33.4%)
  * Among these cases, there were 223 incident cases within 6 months after AHS-2 enrollment
  * Similarly, numbers of incident cases within 12, 18, 24 36, and 48 months are shown below:

```{r incidence_after_enrollment, echo = FALSE}
# How many were diagnosed within 6, 12,...months after enrollment?
c(6, 12, 18, 24, 36, 48) %>% 
  map_df(\(x) 
         tibble(
           `Months after enrollment` = x,
           `N incident CVD cases`    = filter(mi1, CVD_EVER <= qreturndate %m+% months(x)) %>% nrow()
           )
         ) %>% 
  knitr::kable()

```

* For each of the five CVD conditions, the number of incident cases is shown below:

```{r five_cvd_cc, echo = FALSE}

cvd_vars <- c("ISCHEMICHEART_EVER", "STROKE_TIA_EVER", "AMI_EVER", "CHF_EVER", "ATRIAL_FIB_EVER")

# Number of incident CVD cases
cvd_df <- mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), "No", "Yes"))

CreateTableOne(cvd_vars, data = cvd_df) %>% kableone()

```

* Many of these CVD cases have more than one of the 5 conditions. See below for the frequency table of CVD conditions:

```{r n_cvd_cc, echo = FALSE}

# Number of CVDs per subject
mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), 0, 1)) %>% 
  mutate(`Number of CVD conditions` = rowSums(across(all_of(cvd_vars)))) %>% 
  count(`Number of CVD conditions`) %>%
  mutate(Percent = round(n / sum(n) * 100, 2)) %>%
  knitr::kable()

```

## Descriptive tables

### Person-years of follow-up

```{r desc_person_years, echo = FALSE}

# Summary statistics of follow-up years
mi1 %>% 
  summarize(
    `Total person-years`      = sum(fuyear),
    `Follow-up years: Mean`   = mean(fuyear),
    `Follow-up years: Median` = median(fuyear),
    `Follow-up years: SD`     = sd(fuyear)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  knitr::kable(digits = 2)

```

### Age at diagnosis of incident CVD cases 

```{r desc_age_dx, echo = FALSE}

# Age at diagnosis: Mean 78.5 years, Median 78.5 years
mi1 %>% 
  filter(CVD_YN == "Yes") %>% 
  summarize(
    N                          = n(),
    `Age at diagnosis: Mean`   = mean(ageout),
    `Age at diagnosis: Median` = median(ageout),
    `Age at diagnosis: SD`     = sd(ageout)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  mutate(Value = ifelse(Statistic == "N", as.character(Value), sprintf("%.2f", Value))) %>% 
  knitr::kable()
```

### By cases/non-cases

```{r desc_by_case_noncase, echo = FALSE}

# Variables to be included
tablevars <- c("agecat", 
               "bene_age_at_end_2022", 
               "bene_sex_F", 
               "rti_race3", 
               "marital", 
               "educyou", 
               "vegstat", 
               "bmicat", 
               "bmi", 
               "exercise", 
               "sleephrs", 
               "smokecat6", 
               "alccat", 
               "como_depress",
               "como_disab", 
               "como_diabetes", 
               "como_hypert", 
               "como_hyperl", 
               "como_resp", 
               "como_anemia", 
               "como_kidney", 
               "como_hypoth", 
               "como_cancers",
               "egg_freq4",
               "meat_gram_ea",
               "fish_gram_ea",
               "alldairy2_gram_ea",
               "totalveg_gram_ea",
               "fruits_gram_ea",
               "refgrains_gram_ea",
               "whole_mixed_grains_gram_ea",
               "nutsseeds_gram_ea",
               "legumes_gram_ea"
               )

mi1 %>%
  mutate(CVD_YN2 = fct_recode(CVD_YN, "Non-case" = "No", "Case" = "Yes")) %>%
  CreateTableOne(tablevars, strata = "CVD_YN2", data = ., addOverall = TRUE) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()

```

## Cox proportional hazards models

### Egg x meat interaction 




