---
title: "Egg composite CVD study"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "tableone", "survival", "mice")
sapply(pacs, require, character.only = TRUE)

# Read the list of imputed data -------------------------------------------

imputed_data_list <- readRDS("./Data/imputed_data_list.rds")

# First imputed data
mi1 <- imputed_data_list[[1]]

```

## Aim

* Examine associations between egg and meat intake and incident cardiovascular disease (CVD)
* Examine whether egg intake modifies the association between meat intake and CVD
  * i.e., whether egg and meat intake interact in their effect on CVD incidence

## Outcome

* Any of the following incident CVD cases:
  * Ischemic heart disease
  * Stroke/transient ischemic attack
  * Acute myocardial infarction
  * Congestive heart failure (in 27 chronic conditions) or heart failure and non-ischemic heart disease (in 30 chronic conditions)
  * Atrial fibrillation
* For disease definition, see [Chronic Conditions Data Warehouse](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
  * Note that the definitions may differ between 27 chronic conditions (available in 2008-2020 data) and 30 chronic conditions (2021-2022)

## Exposure of interest

* Dietary intake, measured via the AHS baseline questionnaire

* Egg intake frequency
  * Categorized into four groups based on frequency of intake

* Meat intake, gram/day
  * Energy-adjusted
  
* Multiplicative interaction terms between egg intake category and meat intake (g/d) will be assessed in Cox proportional hazards models

## Datasets

* Medicare data, 2008-2022
  * For details regarding Medicare data, see [AHS-2 Medicare Linkage](https://github.com/keijioda/ahs_medicare_linkage/blob/main/summary.md) repository.
  
  * Master Beneficiary Summary File (MBSF), 2008-2022
    * Contains beneficiary characteristics and enrollment information
    
  * Chronic Conditions file (CC), 2008-2022
    * Contains the first occurrence date of [27 (or 30) specific chronic conditions](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
    * Used to identify prevalent/incident cases of CVD and
    * to identify comorbidities, based on ICD-9 and ICD-10 codes
    
  * Both files include n = 46,897 unique subjects across years, after excluding 
    * Gender/DOB mismatch with AHS-2 data
    * Dupulicate beneficiary IDs and SSNs

* AHS-2 baseline data: n = 96,144
  
* After merging Medicare and AHS-2 data, there were n = 43,467 subjects
  * Participants who opted out from the study or those who live outside the U.S were already excluded at this point

## Inclusion/exclusion criteria

* Medicare beneficiaries who did not reach the age of 65 between 2008 and 2022 (e.g., younger beneficiaries with disabilities or end-stage renal disease) were excluded (n = 1238), resulting n = 42,229.

* n = 106 subjects with extreme BMI (<16 or >60), according to AHS questionnaire, were excluded, resulting n = 42,123.

* Unverified dates of deaths
  * Medicare data include a variable (`VALID_DEATH_DT_SW`) indicating whether a beneficiary's day of death has been verified by the Social Security Administration or the Railroad Retirement Board.
  * There were 23 unverified death dates. Excluding these resulted in n = 42,100.

* Prevalent CVD cases
  * At this point of analysis, prevalent cases were defined as those who were diagnosed with any of the five CVD conditions prior to their AHS-2 enrollment
  * There were 3,542 such prevalent cases. Excluding these prevalence resulted in n = 38,558
 
## Multiple imputation

* For AHS-2 baseline data, including food-frequency questionnaire (FFQ), a guided multiple imputation was used to fill missing data ([Fraser & Yan, 2007](https://pubmed.ncbi.nlm.nih.gov/17259903/))
  * Five imputed data sets were generated for subsequent analyses (See the analysis section for more details)
  * For descriptive analysis, we present results from the first imputed data
 
## Incident CVD cases

* Among n = 38,558 subjects, there were 12,890 incident CVD cases (33.4%)
  * Among these cases, there were 223 incident cases within 6 months after AHS-2 enrollment
  * Similarly, numbers of incident cases within 12, 18, 24 36, and 48 months are shown below:

```{r incidence_after_enrollment, echo = FALSE}
# How many were diagnosed within 6, 12,...months after enrollment?
c(6, 12, 18, 24, 36, 48) %>% 
  map_df(\(x) 
         tibble(
           `Months after enrollment` = x,
           `N incident CVD cases`    = filter(mi1, CVD_EVER <= qreturndate %m+% months(x)) %>% nrow()
           )
         ) %>% 
  knitr::kable()

```

* For each of the five CVD conditions, the number of incident cases is shown below:

```{r five_cvd_cc, echo = FALSE}

cvd_vars <- c("ISCHEMICHEART_EVER", "STROKE_TIA_EVER", "AMI_EVER", "CHF_EVER", "ATRIAL_FIB_EVER")

# Number of incident CVD cases
cvd_df <- mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), "No", "Yes"))

CreateTableOne(cvd_vars, data = cvd_df) %>% kableone()

```

* Many of these CVD cases have more than one of the 5 conditions. See below for the frequency table of CVD conditions:

```{r n_cvd_cc, echo = FALSE}

# Number of CVDs per subject
mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), 0, 1)) %>% 
  mutate(`Number of CVD conditions` = rowSums(across(all_of(cvd_vars)))) %>% 
  count(`Number of CVD conditions`) %>%
  mutate(Percent = round(n / sum(n) * 100, 2)) %>%
  knitr::kable()

```

## Descriptive tables

### Person-years of follow-up

```{r desc_person_years, echo = FALSE}

# Summary statistics of follow-up years
mi1 %>% 
  summarize(
    `Total person-years`      = sum(fuyear),
    `Follow-up years: Mean`   = mean(fuyear),
    `Follow-up years: Median` = median(fuyear),
    `Follow-up years: SD`     = sd(fuyear)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  knitr::kable(digits = 2)

```

### Age at diagnosis of incident CVD cases 

```{r desc_age_dx, echo = FALSE}

# Age at diagnosis: Mean 78.5 years, Median 78.5 years
mi1 %>% 
  filter(CVD_YN == "Yes") %>% 
  summarize(
    N                          = n(),
    `Age at diagnosis: Mean`   = mean(ageout),
    `Age at diagnosis: Median` = median(ageout),
    `Age at diagnosis: SD`     = sd(ageout)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  mutate(Value = ifelse(Statistic == "N", as.character(Value), sprintf("%.2f", Value))) %>% 
  knitr::kable()
```

### By cases/non-cases

```{r desc_by_case_noncase, echo = FALSE}

# Variables to be included
tablevars <- c("agecat", 
               "bene_age_at_end_2022", 
               "bene_sex_F", 
               "rti_race3", 
               "marital", 
               "educyou", 
               "vegstat", 
               "bmicat", 
               "bmi", 
               "exercise", 
               "sleephrs", 
               "smokecat6", 
               "alccat", 
               "como_depress",
               "como_disab", 
               "como_diabetes", 
               "como_hypert", 
               "como_hyperl", 
               "como_resp", 
               "como_anemia", 
               "como_kidney", 
               "como_hypoth", 
               "como_cancers",
               "egg_freq4",
               "meat_gram_ea",
               "fish_gram_ea",
               "alldairy2_gram_ea",
               "totalveg_gram_ea",
               "fruits_gram_ea",
               "refgrains_gram_ea",
               "whole_mixed_grains_gram_ea",
               "nutsseeds_gram_ea",
               "legumes_gram_ea"
               )

mi1 %>%
  mutate(CVD_YN2 = fct_recode(CVD_YN, "Non-case" = "No", "Case" = "Yes")) %>%
  CreateTableOne(tablevars, strata = "CVD_YN2", data = ., addOverall = TRUE) %>% 
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()

```

## Cox proportional hazards models

* To examine risk factors associated with incident CVD, we employed the Cox proportional hazards model with attained age as the time scale
  * Age at entry was calculated based on the return date of AHS-2 questionnaire
  * Those who died during the follow-up were censored at the date of death verified in Medicare data
  * Those who were diagnosed with any of the CVD conditions after study enrollment were identified as incident cases and their age at diagnosis was calculated
  * The main exposure variable of interest was the frequency of egg intake.
    * In the AHS-2 baseline questionnaire, the frequency of egg intake was measured in 9 categories
    * Based on its distribution, egg frequency was re-categorized into 4 groups (see the "Descriptive table" section)
  * For meat and other food group variables, their intake was calculated in gram per day and then energy-adjusted by the residual method

* Other than egg frequency (4 levels), meat intake (gram/day, energy-adjusted), and egg x meat interaction, the Cox model includes:
  * Demographics: 
    * From Medicare data: Sex, RTI race
    * From AHS-2 baseline questioannaire: Marital status, educational level
  * Lifestyles: BMI category, exercise, sleep hours, smoking status, alcohol use
  * Dietary intake:
    * Total energy intake (kcal/day)
    * Fish intake (gram/day, energy-adjusted)
    * Dairy intake (gram/day, energy-adjusted)
    * Total vegetable intake (gram/day, energy-adjusted)
    * Fruit intake (gram/day, energy-adjusted)
    * Refined grain intake (gram/day, energy-adjusted)
    * Whole/mixed grain intake (gram/day, energy-adjusted)
    * Nuts/seeds intake (gram/day, energy-adjusted)
    * Legume intake (gram/day, energy-adjusted)
  * Comorbidity (from Medicare data)
    * Depression
    * Functional disabilities (cataract, glaucoma, hip fracture, osteoporosis, rheumatoid arthritis/osteoarthritis)
    * Diabetes
    * Hyperlipidemia
    * Respiratory diseases (COPD, asthma)
    * Anemia
    * Chronic kidney diseases
    * Hypothyroidism
    * Cancers (breast, colorectal, prostate, lung, endometrial)

* The models were run for each imputed data sets, yielding 5 sets of estimated beta coefficients and their variance-covariance matrix
  * These results were combined according to Rubin's rule, and the pooled estimates of HRs and their 95% confidence intervals were produced
  * All analyses were performed on R version 4.6.1

### Hazard ratios for variables other than egg and meat intake

* See the estimated hazard ratios for variables other than egg and meat intake
  * Note that the reference levels are omitted in the table below

```{r cox_model_HR_other, echo = FALSE}

# Cox models --------------------------------------------------------------

# Model 1 -----------------------------------------------------------------

# Demographics, lifestyles, egg intake (as categorical) and kcal
mod1_full_fm <- Surv(agein, ageout, inc_CVD) ~ egg_freq4 + bene_sex_F + rti_race3 + marital +
  educyou2 + bmicat + exercise + sleephrs2 + smokecat6 + alccat + kcal100

# Model 2 -----------------------------------------------------------------

# Adjusting for other foods
# Meat as categorical, other food groups as continuous

# Full model
mod2_full_fm <- update(
  mod1_full_fm, 
  . ~ . + meat_gram_ea100 + fish_gram_ea100 + alldairy2_gram_ea100 + totalveg_gram_ea100 +
    fruits_gram_ea100 + refgrains_gram_ea100 + whole_mixed_grains_gram_ea100 +
    nutsseeds_gram_ea100 + legumes_gram_ea100
)

# Model 3 -----------------------------------------------------------------

# Further adjusting for comorbidity 

# Full model
mod3_full_fm <- update(
  mod2_full_fm, 
  . ~ . + como_depress + como_disab + como_diabetes + como_hyperl  + como_resp + 
    como_anemia + como_kidney + como_hypoth + como_cancers
)

# Checking interactions ---------------------------------------------------

# Based on Model 3
# Check for egg x meat interaction -- Significant p = 0.018
mod4a_full_fm <- update(mod3_full_fm, . ~ . + egg_freq4 * meat_gram_ea100)

fit_mod4a_full <- lapply(imputed_data_list, function(d) {
  coxph(mod4a_full_fm, data = d, method = "efron")
})

mira_fit_mod4a_full <- as.mira(fit_mod4a_full)

# HRs for other variables -------------------------------------------------

# Extract variable names from the formula (preserve model order)
var_names_ordered <- all.vars(mod4a_full_fm)[-(1:3)] %>%   # drop agein, ageout, inc_CVD
  unique()

# Longest-first version, used only for matching (avoids partial-match issues like sleephrs2 vs sleephrs)
var_names_for_match <- var_names_ordered[order(-nchar(var_names_ordered))]

extract_var_level <- function(term, var_names) {
  term <- as.character(term)
  match_idx <- which(startsWith(term, var_names))[1]
  if (is.na(match_idx)) {
    return(tibble(Variable = term, Level = NA_character_))
  }
  v <- var_names[match_idx]
  lvl <- sub(paste0("^", v), "", term)
  tibble(Variable = v, Level = ifelse(lvl == "", NA_character_, lvl))
}

result <- mira_fit_mod4a_full %>% 
  pool() %>% 
  summary(conf.int = TRUE, exp = TRUE, conf.level = .95) %>% 
  select(term, estimate, conf.low, conf.high, p.value) %>% 
  filter(!str_detect(term, "egg|meat")) %>% 
  mutate(
    HR = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
    p.value = ifelse(p.value < 0.0001, "<.0001", sprintf("%.4f", p.value))
  )

var_level <- purrr::map_dfr(as.character(result$term), extract_var_level, var_names = var_names_for_match)

result %>% 
  bind_cols(var_level) %>% 
  select(Variable, Level, HR, p.value) %>% 
  mutate(Level = ifelse(is.na(Level), "", Level)) %>% 
  knitr::kable(row.names = FALSE)

```

### Hazard ratios with egg x meat interaction 

* No egg & no meat as reference

```{r egg_meat_intx_HRs, echo = FALSE}

# HRs with egg x meat interaction -----------------------------------------

# Pool coefficient vector + FULL covariance matrix via Rubin's rules
# (this is what pool() does internally before it collapses to per-term
# variances — we need to stop one step earlier and keep the matrix)
pool_coef_vcov <- function(fit_list) {
  m <- length(fit_list)
  coefs <- lapply(fit_list, coef)
  vcovs <- lapply(fit_list, vcov)
  
  qbar <- Reduce(`+`, coefs) / m                      # pooled point estimate
  ubar <- Reduce(`+`, vcovs) / m                       # within-imputation covariance
  b    <- Reduce(`+`, lapply(coefs, function(q) tcrossprod(q - qbar))) / (m - 1)  # between-imputation covariance
  t_mat <- ubar + (1 + 1 / m) * b                      # total pooled covariance
  
  list(qbar = qbar, vcov = t_mat)
}

# Generalized joint HR: egg level e vs "None", meat = m vs 0, from the SAME
# joint reference cell (None, meat=0) -- includes the meat main effect,
# unlike intx_hr_egg_pooled() which held meat fixed across the comparison.
joint_hr_pooled <- function(fit_list, meat_main_term, egg_main_term = NULL,
                            egg_meat_intx_term = NULL, meat_gram_value) {
  p <- pool_coef_vcov(fit_list)
  qbar <- p$qbar
  V    <- p$vcov
  
  L <- qbar
  L[] <- 0
  L[meat_main_term] <- meat_gram_value
  if (!is.null(egg_main_term))      L[egg_main_term]      <- 1
  if (!is.null(egg_meat_intx_term)) L[egg_meat_intx_term] <- meat_gram_value
  
  est <- qbar %*% L
  se  <- sqrt(t(L) %*% V %*% L)
  lwr <- est - qnorm(0.975) * se
  upr <- est + qnorm(0.975) * se
  c(est, lwr, upr) %>% exp()
}

# Egg terms, including "None" (no main/interaction term -- reference level)
egg_terms <- list(
  "None"   = list(main = NULL, intx = NULL),
  "1-3/mo" = list(main = "egg_freq41-3/mo", intx = "egg_freq41-3/mo:meat_gram_ea100"),
  "1-4/wk" = list(main = "egg_freq41-4/wk", intx = "egg_freq41-4/wk:meat_gram_ea100"),
  "5+/wk"  = list(main = "egg_freq45+/wk",  intx = "egg_freq45+/wk:meat_gram_ea100")
)

# Meat values in grams/day, as in the table -- converted to the model's
# _100 scale (meat_gram_ea100 = meat_gram_ea / 100) inside the call
meat_values_gramday <- c(0, 10, 30, 100)

hr_table <- expand_grid(
  egg_freq4    = names(egg_terms),
  meat_gramday = meat_values_gramday
) %>%
  pmap_dfr(function(egg_freq4, meat_gramday) {
    terms <- egg_terms[[egg_freq4]]
    hr <- joint_hr_pooled(
      fit_mod4a_full,
      meat_main_term     = "meat_gram_ea100",
      egg_main_term      = terms$main,
      egg_meat_intx_term = terms$intx,
      meat_gram_value    = meat_gramday / 100
    )
    data.frame(egg_freq4 = egg_freq4, meat_gramday = meat_gramday,
               HR = hr[1], lwr = hr[2], upr = hr[3])
  }) %>%
  mutate(egg_freq4 = factor(egg_freq4, levels = names(egg_terms)))


hr_table_wide <- hr_table %>%
  mutate(cell = sprintf("%.2f (%.2f, %.2f)", HR, lwr, upr)) %>%
  mutate(`Meat intake (gram/day)` = meat_gramday) %>% 
  select(`Meat intake (gram/day)`, egg_freq4, cell) %>%
  pivot_wider(names_from = egg_freq4, values_from = cell)

hr_table_wide %>% 
  knitr::kable()

```

### Graphical representation of HRs


```{r egg_meat_intx_HR_forest_plot, echo = FALSE, fig.height = 8, fig.width = 9}

# Forest plot
meat_values_gramday <- c(0, 10, 30, 100)   # match what you used to build hr_table

plot_data <- hr_table %>%
  mutate(
    egg_label = recode(as.character(egg_freq4),
                       "None"   = "Egg intake: None",
                       "1-3/mo" = "Egg intake: 1-3 times/month",
                       "1-4/wk" = "Egg intake: 1-4 times/week",
                       "5+/wk"  = "Egg intake: 5+ times/week"),
    egg_label = factor(egg_label, levels = c("Egg intake: None",
                                             "Egg intake: 1-3 times/month",
                                             "Egg intake: 1-4 times/week",
                                             "Egg intake: 5+ times/week")),
    meat_label = factor(paste0("Meat: ", meat_gramday, " g/d"),
                        levels = paste0("Meat: ", rev(meat_values_gramday), " g/d")),
    is_ref  = egg_freq4 == "None" & meat_gramday == 0,
    sig_cat = case_when(
      is_ref  ~ "Reference",
      lwr > 1 ~ "HR > 1, significant",
      upr < 1 ~ "HR < 1, significant",
      TRUE    ~ "Not significant"
    ),
    sig_cat = factor(sig_cat, levels = c("HR > 1, significant", "HR < 1, significant",
                                         "Not significant", "Reference")),
    label = ifelse(is_ref, "1.00 (Ref)", sprintf("%.2f (%.2f, %.2f)", HR, lwr, upr))
  )

x_label_pos <- max(plot_data$upr, na.rm = TRUE) * 1.15   # just past the widest CI bar
x_max       <- x_label_pos * 1.20                          # room for the text column itself

p <- ggplot(plot_data, aes(x = HR, y = meat_label, color = sig_cat, shape = sig_cat)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey60") +
  geom_errorbar(aes(xmin = lwr, xmax = upr), width = 0, linewidth = 0.7,
                orientation = "y", data = filter(plot_data, !is_ref)) +
  geom_point(size = 3) +
  geom_text(aes(x = x_label_pos, label = label), color = "black", hjust = 0, size = 3.6) +
  facet_wrap(vars(egg_label), ncol = 1, scales = "free_y", strip.position = "top") +
  scale_x_log10(breaks = c(0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0),
                limits = c(0.75, x_max),
                expand = expansion(mult = c(0.02, 0.02))) +
  scale_color_manual(values = c("HR > 1, significant" = "#D2691E",
                                 "HR < 1, significant" = "#1B5E56",
                                 "Not significant"     = "grey55",
                                 "Reference"            = "black"),
                     breaks = c("HR > 1, significant", "HR < 1, significant")) +
  scale_shape_manual(values = c("HR > 1, significant" = 16,
                                 "HR < 1, significant" = 16,
                                 "Not significant"     = 16,
                                 "Reference"             = 18),
                     breaks = c("HR > 1, significant", "HR < 1, significant")) +
   labs(x = "Hazard ratio (log scale)", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    strip.placement    = "outside",
    strip.text         = element_text(face = "bold", hjust = 0, size = 13),
    strip.background   = element_rect(fill = "grey93", color = NA),
    panel.spacing      = unit(0.6, "lines"),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    plot.title         = element_text(face = "bold"),
    axis.text.y        = element_text(size = 11)
  ) +
  coord_cartesian(clip = "off")

print(p)
```
