---
title: "Egg composite CVD study"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Required packages
pacs <- c("tidyverse", "tableone", "survival", "mice", "kableExtra", "patchwork")
sapply(pacs, require, character.only = TRUE)

# Read the list of imputed data -------------------------------------------

imputed_data_list <- readRDS("./Data/imputed_data_list.rds")

# First imputed data
mi1 <- imputed_data_list[[1]]

```

## Aim

* Examine associations between egg and meat intake and incident cardiovascular disease (CVD)
* Examine whether egg intake modifies the association between meat intake and CVD
  * i.e., whether egg and meat intake interact in their effect on CVD incidence

## Outcome

* Any of the following incident CVD cases:
  * Ischemic heart disease
  * Stroke/transient ischemic attack
  * Acute myocardial infarction
  * Congestive heart failure (in 27 chronic conditions) or heart failure and non-ischemic heart disease (in 30 chronic conditions)
  * Atrial fibrillation
* For disease definitions, see [Chronic Conditions Data Warehouse](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
  * Note that the definitions may differ between 27 chronic conditions (available in 2008-2020 data) and 30 chronic conditions (2021-2022)

## Exposure of interest

* Dietary intake, measured via the AHS baseline questionnaire

* Egg intake frequency
  * Categorized into four groups based on frequency of intake

* Meat intake, gram/day
  * Energy-adjusted
  
* Multiplicative interaction terms between egg intake category and meat intake (g/d) were assessed in Cox proportional hazards models

## Datasets

* Medicare data, 2008-2022
  * For details regarding Medicare data, see [AHS-2 Medicare Linkage](https://github.com/keijioda/ahs_medicare_linkage/blob/main/summary.md) repository.
  
  * Master Beneficiary Summary File (MBSF), 2008-2022
    * Contains beneficiary characteristics and enrollment information
    
  * Chronic Conditions file (CC), 2008-2022
    * Contains the first occurrence date of [27 (or 30) specific chronic conditions](https://www2.ccwdata.org/web/guest/condition-categories-chronic)
    * Used to identify prevalent/incident cases of CVD and to identify comorbidities, based on ICD-9 and ICD-10 codes
    
  * Both files include n = 46,897 unique subjects across years, after excluding 
    * Gender/DOB mismatch with AHS-2 data
    * Duplicate beneficiary IDs and SSNs

* AHS-2 baseline data: n = 96,144
  
* After merging Medicare and AHS-2 data, there were n = 43,467 subjects
  * Participants who opted out from the study or those who live outside the U.S. were already excluded at this point

## Inclusion/exclusion criteria

* Medicare beneficiaries who did not reach the age of 65 between 2008 and 2022 (e.g., younger beneficiaries with disabilities or end-stage renal disease) were excluded (n = 1238), resulting in n = 42,229.

* n = 106 subjects with extreme BMI (<16 or >60), according to AHS questionnaire, were excluded, resulting n = 42,123.

* Unverified dates of deaths
  * Medicare data include a variable (`VALID_DEATH_DT_SW`) indicating whether a beneficiary's day of death has been verified by the Social Security Administration or the Railroad Retirement Board.
  * There were 23 unverified death dates. Excluding these resulted in n = 42,100.

* Prevalent CVD cases
  * At this point of analysis, prevalent cases were defined as those who were diagnosed with any of the five CVD conditions prior to their AHS-2 enrollment
  * There were 3,542 such prevalent cases. Excluding these prevalen cases resulted in n = 38,558
 
## Multiple imputation

* For AHS-2 baseline data, including food-frequency questionnaire (FFQ), a guided multiple imputation was used to fill in missing data ([Fraser & Yan, 2007](https://pubmed.ncbi.nlm.nih.gov/17259903/))
  * Five imputed data sets were generated for subsequent analyses (See the analysis section for more details)
  * For descriptive analysis, we present results from the first imputed dataset
 
## Incident CVD cases

* Among n = 38,558 subjects, there were 12,890 incident CVD cases (33.4%)
  * Among these cases, there were 223 incident cases within 6 months after AHS-2 enrollment
  * Similarly, numbers of incident cases within 12, 18, 24, 36, and 48 months are shown below:

```{r incidence_after_enrollment, echo = FALSE}
# How many were diagnosed within 6, 12,...months after enrollment?
c(6, 12, 18, 24, 36, 48) %>% 
  map_df(\(x) 
         tibble(
           `Months after enrollment` = x,
           `N incident CVD cases`    = filter(mi1, CVD_EVER <= qreturndate %m+% months(x)) %>% nrow()
           )
         ) %>% 
  knitr::kable()

```

* For each of the five CVD conditions, the number of incident cases is shown below:

```{r five_cvd_cc, echo = FALSE}

cvd_vars <- c("ISCHEMICHEART_EVER", "STROKE_TIA_EVER", "AMI_EVER", "CHF_EVER", "ATRIAL_FIB_EVER")

# Number of incident CVD cases
cvd_df <- mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), "No", "Yes"))

CreateTableOne(cvd_vars, data = cvd_df) %>% kableone()

```

* Many of these CVD cases have more than one of the 5 conditions. See below for the frequency table of CVD conditions:

```{r n_cvd_cc, echo = FALSE}

# Number of CVDs per subject
mi1 %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), 0, 1)) %>% 
  mutate(`Number of CVD conditions` = rowSums(across(all_of(cvd_vars)))) %>% 
  count(`Number of CVD conditions`) %>%
  mutate(Percent = round(n / sum(n) * 100, 2)) %>%
  knitr::kable()

```

## Descriptive tables

### Person-years of follow-up

```{r desc_person_years, echo = FALSE}

# Summary statistics of follow-up years
mi1 %>% 
  summarize(
    `Total person-years`      = sum(fuyear),
    `Follow-up years: Mean`   = mean(fuyear),
    `Follow-up years: Median` = median(fuyear),
    `Follow-up years: SD`     = sd(fuyear)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  knitr::kable(digits = 2)

```

### Age at diagnosis of incident CVD cases 

```{r desc_age_dx, echo = FALSE}

# Age at diagnosis: Mean 78.5 years, Median 78.5 years
mi1 %>% 
  filter(CVD_YN == "Yes") %>% 
  summarize(
    N                          = n(),
    `Age at diagnosis: Mean`   = mean(ageout),
    `Age at diagnosis: Median` = median(ageout),
    `Age at diagnosis: SD`     = sd(ageout)
  ) %>% 
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>% 
  mutate(Value = ifelse(Statistic == "N", as.character(Value), sprintf("%.2f", Value))) %>% 
  knitr::kable()
```

### By cases/non-cases

```{r desc_by_case_noncase, echo = FALSE}

# Variables to be included
tablevars <- c("agecat", 
               "bene_age_at_end_2022", 
               "bene_sex_F", 
               "rti_race3", 
               "marital", 
               "educyou", 
               "vegstat", 
               "bmicat", 
               "bmi", 
               "exercise", 
               "sleephrs", 
               "smokecat6", 
               "alccat", 
               "como_depress",
               "como_disab", 
               "como_diabetes", 
               "como_hypert", 
               "como_hyperl", 
               "como_resp", 
               "como_anemia", 
               "como_kidney", 
               "como_hypoth", 
               "como_cancers",
               "egg_freq4",
               "meat_gram_ea",
               "fish_gram_ea",
               "alldairy2_gram_ea",
               "totalveg_gram_ea",
               "fruits_gram_ea",
               "refgrains_gram_ea",
               "whole_mixed_grains_gram_ea",
               "nutsseeds_gram_ea",
               "legumes_gram_ea"
               )

mi1 %>%
  mutate(CVD_YN2 = fct_recode(CVD_YN, "Non-case" = "No", "Case" = "Yes")) %>%
  CreateTableOne(tablevars, strata = "CVD_YN2", data = ., addOverall = TRUE) %>% 
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()

```

## Cox proportional hazards models

* To examine risk factors associated with incident CVD, we employed the Cox proportional hazards model with attained age as the time scale
  * Age at entry was calculated based on the return date of AHS-2 questionnaire
  * Those who died during the follow-up were censored at the date of death verified in Medicare data
  * Those who were diagnosed with any of the CVD conditions after study enrollment were identified as incident cases and their age at diagnosis was calculated
  * The main exposure variable of interest was the frequency of egg intake.
    * In the AHS-2 baseline questionnaire, the frequency of egg intake was measured in 9 categories
    * Based on its distribution, egg frequency was re-categorized into 4 groups (see the "Descriptive table" section)
  * For meat and other food group variables, their intake was calculated in gram per day and then energy-adjusted by the residual method

* Other than egg frequency (4 levels), meat intake (gram/day, energy-adjusted), and egg x meat interaction, the Cox model includes:
  * Demographics: 
    * From Medicare data: Sex, RTI race
    * From AHS-2 baseline questionnaire: Marital status, educational level
  * Lifestyles: BMI category, exercise, sleep hours, smoking status, alcohol use
  * Dietary intake:
    * Total energy intake (kcal/day)
    * Fish intake (gram/day, energy-adjusted)
    * Dairy intake (gram/day, energy-adjusted)
    * Total vegetable intake (gram/day, energy-adjusted)
    * Fruit intake (gram/day, energy-adjusted)
    * Refined grain intake (gram/day, energy-adjusted)
    * Whole/mixed grain intake (gram/day, energy-adjusted)
    * Nuts/seeds intake (gram/day, energy-adjusted)
    * Legume intake (gram/day, energy-adjusted)
  * Comorbidity (from Medicare data)
    * Depression
    * Functional disabilities (cataract, glaucoma, hip fracture, osteoporosis, rheumatoid arthritis/osteoarthritis)
    * Diabetes
    * Hyperlipidemia
    * Respiratory diseases (COPD, asthma)
    * Anemia
    * Chronic kidney diseases
    * Hypothyroidism
    * Cancers (breast, colorectal, prostate, lung, endometrial)

* The models were run for each of the imputed data sets, yielding 5 sets of estimated beta coefficients and their variance-covariance matrix
  * These results were combined according to Rubin's rules, and the pooled estimates of HRs and their 95% confidence intervals were produced
  
* The proportional hazards assumption was assessed by plotting scaled Schoenfeld residuals against time for each covariate. The residuals scattered randomly around zero with no discernible trend over time, supporting the assumption (see results below)

* All analyses were performed in R version 4.6.1

### Hazard ratios for variables other than egg and meat intake

* In the Cox model, the interaction between egg frequency and meat intake was statistically significant (p = 0.018)

* See the estimated hazard ratios for variables other than egg and meat intake
  * Note that the reference levels are omitted in the table below

```{r cox_model_HR_other, echo = FALSE}

# Cox models --------------------------------------------------------------

# Model 1 -----------------------------------------------------------------

# Demographics, lifestyles, egg intake (as categorical) and kcal
mod1_full_fm <- Surv(agein, ageout, inc_CVD) ~ egg_freq4 + bene_sex_F + rti_race3 + marital +
  educyou2 + bmicat + exercise + sleephrs2 + smokecat6 + alccat + kcal100

# Model 2 -----------------------------------------------------------------

# Adjusting for other foods
# Meat as categorical, other food groups as continuous

# Full model
mod2_full_fm <- update(
  mod1_full_fm, 
  . ~ . + meat_gram_ea100 + fish_gram_ea100 + alldairy2_gram_ea100 + totalveg_gram_ea100 +
    fruits_gram_ea100 + refgrains_gram_ea100 + whole_mixed_grains_gram_ea100 +
    nutsseeds_gram_ea100 + legumes_gram_ea100
)

# Model 3 -----------------------------------------------------------------

# Further adjusting for comorbidity 

# Full model
mod3_full_fm <- update(
  mod2_full_fm, 
  . ~ . + como_depress + como_disab + como_diabetes + como_hyperl  + como_resp + 
    como_anemia + como_kidney + como_hypoth + como_cancers
)

# Checking interactions ---------------------------------------------------

# Based on Model 3
# Check for egg x meat interaction -- Significant p = 0.018
mod4a_full_fm <- update(mod3_full_fm, . ~ . + egg_freq4 * meat_gram_ea100)

fit_mod4a_full <- lapply(imputed_data_list, function(d) {
  coxph(mod4a_full_fm, data = d, method = "efron")
})

mira_fit_mod4a_full <- as.mira(fit_mod4a_full)

# HRs for other variables -------------------------------------------------

# Extract variable names from the formula (preserve model order)
var_names_ordered <- all.vars(mod4a_full_fm)[-(1:3)] %>%   # drop agein, ageout, inc_CVD
  unique()

# Longest-first version, used only for matching (avoids partial-match issues like sleephrs2 vs sleephrs)
var_names_for_match <- var_names_ordered[order(-nchar(var_names_ordered))]

extract_var_level <- function(term, var_names) {
  term <- as.character(term)
  match_idx <- which(startsWith(term, var_names))[1]
  if (is.na(match_idx)) {
    return(tibble(Variable = term, Level = NA_character_))
  }
  v <- var_names[match_idx]
  lvl <- sub(paste0("^", v), "", term)
  tibble(Variable = v, Level = ifelse(lvl == "", NA_character_, lvl))
}

result <- mira_fit_mod4a_full %>% 
  pool() %>% 
  summary(conf.int = TRUE, exp = TRUE, conf.level = .95) %>% 
  select(term, estimate, conf.low, conf.high, p.value) %>% 
  filter(!str_detect(term, "egg|meat")) %>% 
  mutate(
    HR = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
    p.value = ifelse(p.value < 0.0001, "<.0001", sprintf("%.4f", p.value))
  )

var_level <- purrr::map_dfr(as.character(result$term), extract_var_level, var_names = var_names_for_match)

result %>% 
  bind_cols(var_level) %>% 
  select(Variable, Level, HR, p.value) %>% 
  mutate(Level = ifelse(is.na(Level), "", Level)) %>% 
  knitr::kable(row.names = FALSE)

```

### Hazard ratios for the joint effect of egg and meat intake 

* One way to illustrate the negative interaction between egg and meat intake is to calculate hazard ratios for each egg intake group at representative levels of meat intake. We chose meat intakes of 0, 10, 30, and 100 grams/day and estimated HRs for all egg × meat combinations, using "no egg intake and no meat intake" as the reference group. These HRs reflect the joint risk of egg and meat intake combined, relative to this "healthiest" reference group (no eggs, no meat).
  * The HRs are shown in the table below
  * Trend p-values for egg frequency, calculated at each level of meat intake, are shown in the last column of the table
  * Trend p-values for meat intake, calculated within each egg frequency group, are shown in the last row of the table
  
```{r egg_meat_intx_HRs, echo = FALSE}

# HRs with egg x meat interaction -----------------------------------------

# Pool the results
pooled <- pool(mira_fit_mod4a_full)

# Pool coefficient vector + FULL covariance matrix via Rubin's rules
# (this is what pool() does internally before it collapses to per-term
# variances — we need to stop one step earlier and keep the matrix)
pool_coef_vcov <- function(fit_list) {
  m <- length(fit_list)
  coefs <- lapply(fit_list, coef)
  vcovs <- lapply(fit_list, vcov)
  
  qbar <- Reduce(`+`, coefs) / m                      # pooled point estimate
  ubar <- Reduce(`+`, vcovs) / m                       # within-imputation covariance
  b    <- Reduce(`+`, lapply(coefs, function(q) tcrossprod(q - qbar))) / (m - 1)  # between-imputation covariance
  t_mat <- ubar + (1 + 1 / m) * b                      # total pooled covariance
  
  list(qbar = qbar, vcov = t_mat)
}

# Generalized joint HR: egg level e vs "None", meat = m vs 0, from the SAME
# joint reference cell (None, meat=0) -- includes the meat main effect,
# unlike intx_hr_egg_pooled() which held meat fixed across the comparison.
joint_hr_pooled <- function(fit_list, meat_main_term, egg_main_term = NULL,
                            egg_meat_intx_term = NULL, meat_gram_value) {
  p <- pool_coef_vcov(fit_list)
  qbar <- p$qbar
  V    <- p$vcov
  
  L <- qbar
  L[] <- 0
  L[meat_main_term] <- meat_gram_value
  if (!is.null(egg_main_term))      L[egg_main_term]      <- 1
  if (!is.null(egg_meat_intx_term)) L[egg_meat_intx_term] <- meat_gram_value
  
  est <- qbar %*% L
  se  <- sqrt(t(L) %*% V %*% L)
  lwr <- est - qnorm(0.975) * se
  upr <- est + qnorm(0.975) * se
  c(est, lwr, upr) %>% exp()
}

# Egg terms, including "None" (no main/interaction term -- reference level)
egg_terms <- list(
  "None"   = list(main = NULL, intx = NULL),
  "1-3/mo" = list(main = "egg_freq41-3/mo", intx = "egg_freq41-3/mo:meat_gram_ea100"),
  "1-4/wk" = list(main = "egg_freq41-4/wk", intx = "egg_freq41-4/wk:meat_gram_ea100"),
  "5+/wk"  = list(main = "egg_freq45+/wk",  intx = "egg_freq45+/wk:meat_gram_ea100")
)

# Meat values in grams/day, as in the table -- converted to the model's
# _100 scale (meat_gram_ea100 = meat_gram_ea / 100) inside the call
meat_values_gramday <- c(0, 10, 30, 100)

hr_table <- expand_grid(
  egg_freq4    = names(egg_terms),
  meat_gramday = meat_values_gramday
) %>%
  pmap_dfr(function(egg_freq4, meat_gramday) {
    terms <- egg_terms[[egg_freq4]]
    hr <- joint_hr_pooled(
      fit_mod4a_full,
      meat_main_term     = "meat_gram_ea100",
      egg_main_term      = terms$main,
      egg_meat_intx_term = terms$intx,
      meat_gram_value    = meat_gramday / 100
    )
    data.frame(egg_freq4 = egg_freq4, meat_gramday = meat_gramday,
               HR = hr[1], lwr = hr[2], upr = hr[3])
  }) %>%
  mutate(egg_freq4 = factor(egg_freq4, levels = names(egg_terms)))

hr_table_wide <- hr_table %>%
  mutate(
    is_ref = egg_freq4 == "None" & meat_gramday == 0,
    cell   = ifelse(is_ref, "1.00 (Ref)", sprintf("%.2f (%.2f, %.2f)", HR, lwr, upr))
  ) %>%
  select(meat_gramday, egg_freq4, cell) %>%
  pivot_wider(names_from = egg_freq4, values_from = cell)

# Incorporate trend-p into the table
# Function for trend p-value
get_wald_p_pooled <- function(fit_list, coef_names, coef_wt) {
  p    <- pool_coef_vcov(fit_list)
  qbar <- p$qbar
  V    <- p$vcov
  
  # Safety check: every requested name must exist in the pooled coefficient
  # vector, or L[coef_names] <- coef_wt would silently create NA entries
  # instead of erroring.
  missing_terms <- setdiff(coef_names, names(qbar))
  if (length(missing_terms) > 0) {
    stop("Term(s) not found in model coefficients: ",
         paste(missing_terms, collapse = ", "))
  }
  
  L <- qbar
  L[] <- 0
  L[coef_names] <- coef_wt
  
  est    <- qbar %*% L
  se     <- sqrt(t(L) %*% V %*% L)
  wald_p <- as.numeric(2 * (1 - pnorm(abs(est / se))))
  return(wald_p)
}

# Collect the meat trend p-values, one per egg group (columns)
p_trend_meat <- c(
  "None"   = summary(pooled) %>% filter(term == "meat_gram_ea100") %>% pull(p.value),
  "1-3/mo" = get_wald_p_pooled(fit_mod4a_full,
                               coef_names = c("meat_gram_ea100", "egg_freq41-3/mo:meat_gram_ea100"),
                               coef_wt    = c(1, 1)),
  "1-4/wk" = get_wald_p_pooled(fit_mod4a_full,
                               coef_names = c("meat_gram_ea100", "egg_freq41-4/wk:meat_gram_ea100"),
                               coef_wt    = c(1, 1)),
  "5+/wk"  = get_wald_p_pooled(fit_mod4a_full,
                               coef_names = c("meat_gram_ea100", "egg_freq45+/wk:meat_gram_ea100"),
                               coef_wt    = c(1, 1))
)

# Replace egg_freq4 with its numeric values, 1, 2, 3, 4
mod4b_full_fm <- update(
  mod4a_full_fm, 
  . ~ . - egg_freq4 - egg_freq4:meat_gram_ea100 + as.numeric(egg_freq4) + as.numeric(egg_freq4):meat_gram_ea100
  )

fit_mod4b_full <- lapply(imputed_data_list, function(d) {
  coxph(mod4b_full_fm, data = d, method = "efron")
})

# Collect the egg trend p-values, one per meat level (rows)
p_trend_egg <- c(
  "0"   = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 0)),
  "10"  = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 10 / 100)),
  "30"  = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 30 / 100)),
  "100" = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 100 / 100))
)

# Helper to format p-values consistently
fmt_p <- function(p) ifelse(p < 0.0001, "<0.0001", sprintf("%.4f", p))

# Add the right-margin column (egg trend p, one per meat row)
hr_table_wide <- hr_table_wide %>%
  mutate(`P-trend (egg)` = fmt_p(p_trend_egg[as.character(meat_gramday)]))

# Add the bottom-margin row (meat trend p, one per egg column)
bottom_row <- setNames(
  as.list(c(
    NA,
    fmt_p(p_trend_meat[c("None", "1-3/mo", "1-4/wk", "5+/wk")]),
    ""
  )),
  names(hr_table_wide)
) %>% as_tibble()

hr_table_final <- hr_table_wide %>%
  mutate(meat_gramday = as.character(meat_gramday)) %>%
  bind_rows(bottom_row) %>%
  mutate(meat_gramday = ifelse(is.na(meat_gramday), "P-trend (meat)", meat_gramday))

hr_table_final %>%
  rename(`Meat intake (g/day)` = meat_gramday) %>%
  kable(
    col.names = c("Meat intake (g/day)", "None", "1-3/mo", "1-4/wk", "5+/wk", "P-trend (egg)"),
    align = "lccccc"
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Egg frequency" = 4, " " = 1))

```

* Shown below is a forest plot depicting the HRs for egg × meat combinations:

```{r egg_meat_intx_HR_forest_plot, echo = FALSE, fig.height = 8, fig.width = 9}

# Forest plot
meat_values_gramday <- c(0, 10, 30, 100)   # match what you used to build hr_table

plot_data <- hr_table %>%
  mutate(
    egg_label = recode(as.character(egg_freq4),
                       "None"   = "Egg intake: None",
                       "1-3/mo" = "Egg intake: 1-3 times/month",
                       "1-4/wk" = "Egg intake: 1-4 times/week",
                       "5+/wk"  = "Egg intake: 5+ times/week"),
    egg_label = factor(egg_label, levels = c("Egg intake: None",
                                             "Egg intake: 1-3 times/month",
                                             "Egg intake: 1-4 times/week",
                                             "Egg intake: 5+ times/week")),
    meat_label = factor(paste0("Meat: ", meat_gramday, " g/d"),
                        levels = paste0("Meat: ", rev(meat_values_gramday), " g/d")),
    is_ref  = egg_freq4 == "None" & meat_gramday == 0,
    sig_cat = case_when(
      is_ref  ~ "Reference",
      lwr > 1 ~ "HR > 1, significant",
      upr < 1 ~ "HR < 1, significant",
      TRUE    ~ "Not significant"
    ),
    sig_cat = factor(sig_cat, levels = c("HR > 1, significant", "HR < 1, significant",
                                         "Not significant", "Reference")),
    label = ifelse(is_ref, "1.00 (Ref)", sprintf("%.2f (%.2f, %.2f)", HR, lwr, upr))
  )

x_label_pos <- max(plot_data$upr, na.rm = TRUE) * 1.15   # just past the widest CI bar
x_max       <- x_label_pos * 1.20                          # room for the text column itself

p <- ggplot(plot_data, aes(x = HR, y = meat_label, color = sig_cat, shape = sig_cat)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey60") +
  geom_errorbar(aes(xmin = lwr, xmax = upr), width = 0, linewidth = 0.7,
                orientation = "y", data = filter(plot_data, !is_ref)) +
  geom_point(size = 3) +
  geom_text(aes(x = x_label_pos, label = label), color = "black", hjust = 0, size = 3.6) +
  facet_wrap(vars(egg_label), ncol = 1, scales = "free_y", strip.position = "top") +
  scale_x_log10(breaks = c(0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0),
                limits = c(0.75, x_max),
                expand = expansion(mult = c(0.02, 0.02))) +
  scale_color_manual(values = c("HR > 1, significant" = "#D2691E",
                                 "HR < 1, significant" = "#1B5E56",
                                 "Not significant"     = "grey55",
                                 "Reference"            = "black"),
                     breaks = c("HR > 1, significant", "HR < 1, significant")) +
  scale_shape_manual(values = c("HR > 1, significant" = 16,
                                 "HR < 1, significant" = 16,
                                 "Not significant"     = 16,
                                 "Reference"             = 18),
                     breaks = c("HR > 1, significant", "HR < 1, significant")) +
   labs(x = "Hazard ratio (log scale)", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    strip.placement    = "outside",
    strip.text         = element_text(face = "bold", hjust = 0, size = 13),
    strip.background   = element_rect(fill = "grey93", color = NA),
    panel.spacing      = unit(0.6, "lines"),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    plot.title         = element_text(face = "bold"),
    axis.text.y        = element_text(size = 11)
  ) +
  coord_cartesian(clip = "off")

print(p)
```

### Hazard ratios for egg frequency estimated at different meat intake level

* Another way to illustrate the egg x meat interaction, and one that isolates the egg effect more directly, is to calculate hazard ratios for egg frequency at each level of meat intake. For meat intakes of 0, 10, 30, and 100 grams/day, we separately estimated HRs for egg frequency groups, using "no egg intake" as the reference at each meat level. These HRs show how the association between egg intake and CVD risk is modified by meat intake


```{r egg_HRs_with_meat_intx, echo = FALSE}

# HRs for egg freq at meat intake levels ----------------------------------

# Function to estimate pooled HR for egg with interaction
intx_hr_egg_pooled <- function(fit_list, egg_main_term, egg_meat_intx_term, meat_gram_value) {
  p    <- pool_coef_vcov(fit_list)
  qbar <- p$qbar
  V    <- p$vcov
  
  # Safety check, same pattern as get_wald_p_pooled()
  needed <- c(egg_main_term, egg_meat_intx_term)
  missing_terms <- setdiff(needed, names(qbar))
  if (length(missing_terms) > 0) {
    stop("Term(s) not found in model coefficients: ",
         paste(missing_terms, collapse = ", "))
  }
  
  L <- qbar
  L[] <- 0
  L[egg_main_term]      <- 1
  L[egg_meat_intx_term] <- meat_gram_value
  
  est <- qbar %*% L
  se  <- sqrt(t(L) %*% V %*% L)
  lwr <- est - qnorm(0.975) * se
  upr <- est + qnorm(0.975) * se
  c(est, lwr, upr) %>% exp()
}

egg_labels <- c("1-3/mo", "1-4/wk", "5+/wk")
egg_main_terms <- c("egg_freq41-3/mo", "egg_freq41-4/wk", "egg_freq45+/wk")
egg_intx_terms <- c("egg_freq41-3/mo:meat_gram_ea100",
                    "egg_freq41-4/wk:meat_gram_ea100",
                    "egg_freq45+/wk:meat_gram_ea100")

# Helper to build the HR table for one meat value, looping over all 3 egg terms
egg_hr_at_meat <- function(fit_list, meat_gramday) {
  purrr::pmap_dfr(
    list(egg_main_terms, egg_intx_terms, egg_labels),
    function(main_term, intx_term, label) {
      hr <- intx_hr_egg_pooled(fit_list, main_term, intx_term, meat_gramday / 100)
      tibble(Meat = paste0(meat_gramday, " g/d"), Egg = label,
             HR = hr[1], Lower = hr[2], Upper = hr[3])
    }
  )
}

meat_values_gramday <- c(0, 10, 30, 100)
egg_hr_table <- purrr::map_dfr(meat_values_gramday, ~ egg_hr_at_meat(fit_mod4a_full, .x))


egg_hr_wide <- egg_hr_table %>%
  mutate(
    meat_gramday = as.numeric(sub(" g/d", "", Meat)),
    cell = sprintf("%.2f (%.2f, %.2f)", HR, Lower, Upper)
  ) %>%
  select(meat_gramday, Egg, cell) %>%
  pivot_wider(names_from = Egg, values_from = cell) %>%
  mutate(None = "1.00 (Ref)") %>%
  select(meat_gramday, None, `1-3/mo`, `1-4/wk`, `5+/wk`) %>%
  arrange(meat_gramday) %>%
  mutate(`P-trend` = fmt_p(p_trend_egg[as.character(meat_gramday)]))

egg_hr_wide %>%
  kable(
    col.names = c("Meat intake (gram/day)", "None", "1-3 times/month",
                  "1-4 times/week", "5+ times/week", "P-trend"),
    align = "lccccc"
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Egg frequency" = 4, " " = 1))

```

```{r egg_HR_by_meat_line_plot, echo = FALSE, fig.height = 3, fig.width = 10}

# Add "None" reference rows (HR = 1, no CI) for each meat level
none_rows <- tibble(
  Meat = paste0(c(0, 10, 30, 100), " g/d"),
  Egg  = "None",
  HR = 1, Lower = NA, Upper = NA,
  meat_gramday = c(0, 10, 30, 100)
)

egg_hr_plot_data <- egg_hr_table %>%
  mutate(meat_gramday = as.numeric(sub(" g/d", "", Meat))) %>%
  bind_rows(none_rows) %>%
  mutate(Egg = factor(Egg, levels = c("None", "1-3/mo", "1-4/wk", "5+/wk")))

ggplot(egg_hr_plot_data, aes(x = Egg, y = HR)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point(aes(shape = Egg == "None", color = Egg == "None"),
             size = 3, show.legend = FALSE) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2) +
  scale_shape_manual(values = c("TRUE" = 18, "FALSE" = 16)) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  scale_x_discrete(labels = c("None" = "None\n(Ref)", "1-3/mo" = "1-3/mo",
                              "1-4/wk" = "1-4/wk", "5+/wk" = "5+/wk")) +
  scale_y_log10(breaks = c(0.6, 0.7, 0.8, 0.9, 1, 1.5, 2)) +
  facet_wrap(~ factor(meat_gramday, levels = c(0, 10, 30, 100),
                      labels = paste0("Meat: ", c(0, 10, 30, 100), " g/d")),
             nrow = 1) +
  labs(x = "Egg intake frequency", y = "Hazard ratio (log scale)") +
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "grey85", color = "grey50"),
    strip.text = element_text(face = "bold")
  )
```


## Substitution analysis

* Substitution analysis was conducted to estimate the isocaloric effect of replacing another dietary component with egg intake, holding total energy intake constant, rather than assuming increased egg consumption occurs independent of overall diet composition
  * Substitutions were modeled for two different comparator foods, legumes and nuts/seeds, to assess whether the egg–meat interaction on CVD risk was sensitive to the specific food displaced
  * The substitution amount for each comparator food was defined as the quantity isocaloric with one egg, corresponding to approximately 60 g of legumes and 14 g of nuts/seeds based on their respective energy densities
    * These gram-equivalents were then scaled by the frequency of egg consumption within each exposure category to obtain a daily gram-equivalent quantity of the comparator food displaced.

* Substitution HRs for egg intake across levels of meat consumption were similar regardless of the specific comparator food (legumes or nuts/seeds), suggesting that the observed interaction between egg and meat intake on CVD risk is not sensitive to the choice of displaced dietary component.


```{r substitute_egg_with_legume, echo = FALSE}

# Substitution analysis ---------------------------------------------------

# Function to extract betas and vcov(beta)
get_pooled_estimates <- function(fit_mira) {
  pooled_summary <- fit_mira %>% pool() %>% summary(conf.int = TRUE)
  bhat <- pooled_summary$estimate
  names(bhat) <- pooled_summary$term
  
  fit_analyses <- fit_mira$analyses
  m <- length(fit_analyses)
  qhats <- sapply(fit_analyses, coef)
  qbar <- rowMeans(qhats)
  vw <- Reduce("+", lapply(fit_analyses, vcov)) / m
  vb <- (1 / (m - 1)) * (qhats - qbar) %*% t(qhats - qbar)
  vt <- vw + (1 + 1 / m) * vb
  dimnames(vt) <- list(names(bhat), names(bhat))
  
  list(bhat = bhat, vt = vt)
}

est <- get_pooled_estimates(fit_mod4a_full %>% as.mira())
bhat <- est$bhat
vt   <- est$vt

# loc    : character vector of term names entering the linear combination
# weight : numeric vector of same length, the multiplier for each term
get_est_beta_substitute <- function(bhat, vt, loc, weight) {
  a <- weight
  b <- bhat[loc]
  est_cbeta <- sum(a * b)
  
  v <- vt[loc, loc]
  subs_b_var <- as.numeric(t(a) %*% v %*% a)
  
  ci <- est_cbeta + qnorm(c(.025, .975)) * sqrt(subs_b_var)
  out <- exp(c(est = est_cbeta, lower = ci[1], upper = ci[2]))
  return(out)
}

# Evaluate at a grid of meat levels
meat_grid <- c(0, 10, 30, 100) / 100

# Names for egg terms and their interaction with meat
egg_terms <- c("egg_freq41-3/mo", "egg_freq41-4/wk", "egg_freq45+/wk")
int_terms <- paste0(egg_terms, ":meat_gram_ea100")

## Substitution with legumes ----------------------------------------------

# calorie-equivalent for each category's midpoint frequency (eggs/month -> legume g/day)
egg_freq <- c(2/30, 2.5/7, 1) 
gram_equivs <- 60 * egg_freq / 100

all_results <- lapply(seq_along(egg_terms), function(i) {
  loc <- c(egg_terms[i], int_terms[i], "legumes_gram_ea100")
  t(sapply(meat_grid, function(m_val) {
    weight <- c(1, m_val, -gram_equivs[i])
    get_est_beta_substitute(bhat, vt, loc, weight)
  }))
})
names(all_results) <- egg_terms

# Format "est (lower, upper)" strings, 2 decimals
fmt_hr <- function(mat) {
  sprintf("%.2f (%.2f, %.2f)", mat[, "est"], mat[, "lower"], mat[, "upper"])
}

# convert to actual grams/day for the table
meat_gramday <- meat_grid * 100

# Wide format
subs_egg_for_legumes_hr_wide <- tibble(
  meat_gramday = meat_gramday,
  None = "1.00 (Ref)",
  `1-3/mo` = fmt_hr(all_results[["egg_freq41-3/mo"]]),
  `1-4/wk` = fmt_hr(all_results[["egg_freq41-4/wk"]]),
  `5+/wk`  = fmt_hr(all_results[["egg_freq45+/wk"]])
)

subs_egg_for_legumes_hr_wide %>% 
  rename(`Meat intake (g/day)` = meat_gramday) %>%
  kable(
    col.names = c("Meat intake (g/day)", "None", "1-3/mo", "1-4/wk", "5+/wk"),
    align = "lcccc"
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Egg frequency (substituted for legumes)" = 4))

```

```{r substitute_egg_with_nuts, echo = FALSE}

## Substitution with nuts/seeds -------------------------------------------

# calorie-equivalent for each category's midpoint frequency (eggs/month -> nuts g/day)
egg_freq <- c(2/30, 2.5/7, 1) 
gram_equivs <- 14 * egg_freq / 100

all_results <- lapply(seq_along(egg_terms), function(i) {
  loc <- c(egg_terms[i], int_terms[i], "nutsseeds_gram_ea100")
  t(sapply(meat_grid, function(m_val) {
    weight <- c(1, m_val, -gram_equivs[i])
    get_est_beta_substitute(bhat, vt, loc, weight)
  }))
})
names(all_results) <- egg_terms

# Format "est (lower, upper)" strings, 2 decimals
fmt_hr <- function(mat) {
  sprintf("%.2f (%.2f, %.2f)", mat[, "est"], mat[, "lower"], mat[, "upper"])
}

# convert to actual grams/day for the table
meat_gramday <- meat_grid * 100

# Wide format
subs_egg_for_nuts_hr_wide <- tibble(
  meat_gramday = meat_gramday,
  None = "1.00 (Ref)",
  `1-3/mo` = fmt_hr(all_results[["egg_freq41-3/mo"]]),
  `1-4/wk` = fmt_hr(all_results[["egg_freq41-4/wk"]]),
  `5+/wk`  = fmt_hr(all_results[["egg_freq45+/wk"]])
)

subs_egg_for_nuts_hr_wide %>% 
  rename(`Meat intake (g/day)` = meat_gramday) %>%
  kable(
    col.names = c("Meat intake (g/day)", "None", "1-3/mo", "1-4/wk", "5+/wk"),
    align = "lcccc"
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Egg frequency (substituted for nuts/seeds)" = 4))

```

## Sensitivity analysis

* A sensitivity analysis was also conducted as follows:
  * We applied a two-year lag window following study entry, reclassifying subjects flagged with any of the five CVD conditions within that window from incident to prevalent and excluding them from the at-risk population.
  * To avoid immortal time bias introduced by this post-entry exclusion, the time origin of the Cox model was shifted to two years after study entry.

* In this sensitivity analysis, the interaction between egg frequency and meat intake remained statistically significant (p = 0.035).
  * Estimated HRs for egg frequency at the same levels of meat intake are shown below. The results were similar to those from the main analysis.

```{r sensitivity_analysis, echo = FALSE}

# Read the list of imputed data -------------------------------------------

imputed_data_list_sens <- readRDS("./Data/imputed_data_list_prev_2yrs.rds")

# Cox models --------------------------------------------------------------

# Model 1 -----------------------------------------------------------------

# Demographics, lifestyles, egg intake (as categorical) and kcal

# Full model
mod1_full_fm <- Surv(agein + 2, ageout, inc_CVD) ~ egg_freq4 + bene_sex_F + rti_race3 + marital +
  educyou2 + bmicat + exercise + sleephrs2 + smokecat6 + alccat + kcal100

# Model 2 -----------------------------------------------------------------

# Adjusting for other foods
# Meat as categorical, other food groups as continuous

# Full model
mod2_full_fm <- update(
  mod1_full_fm, 
  . ~ . + meat_gram_ea100 + fish_gram_ea100 + alldairy2_gram_ea100 + totalveg_gram_ea100 +
    fruits_gram_ea100 + refgrains_gram_ea100 + whole_mixed_grains_gram_ea100 +
    nutsseeds_gram_ea100 + legumes_gram_ea100
)

# Model 3 -----------------------------------------------------------------

# Further adjusting for comorbidity 

# Full model
mod3_full_fm <- update(
  mod2_full_fm, 
  . ~ . + como_depress + como_disab + como_diabetes + como_hyperl  + como_resp + 
    como_anemia + como_kidney + como_hypoth + como_cancers
)

# Fit both models
fit_mod3_full <- lapply(imputed_data_list_sens, function(d) {
  coxph(mod3_full_fm, data = d, method = "efron")
})

# Convert to mira
mira_fit_mod3_full <- as.mira(fit_mod3_full)

# Checking interactions ---------------------------------------------------

# Based on Model 3
# Check for egg x meat interaction -- Significant p = 0.0350
mod4a_full_fm <- update(mod3_full_fm, . ~ . + egg_freq4 * meat_gram_ea100)

fit_mod4a_full <- lapply(imputed_data_list_sens, function(d) {
  coxph(mod4a_full_fm, data = d, method = "efron")
})

mira_fit_mod4a_full <- as.mira(fit_mod4a_full)

# Pool the results
pooled <- pool(mira_fit_mod4a_full)

# P-trend for egg intake --------------------------------------------------

# Replace egg_freq4 with its numeric values, 1, 2, 3, 4
mod4b_full_fm <- update(
  mod4a_full_fm, 
  . ~ . - egg_freq4 - egg_freq4:meat_gram_ea100 + as.numeric(egg_freq4) + as.numeric(egg_freq4):meat_gram_ea100
  )

fit_mod4b_full <- lapply(imputed_data_list_sens, function(d) {
  coxph(mod4b_full_fm, data = d, method = "efron")
})

# Collect the egg trend p-values, one per meat level (rows)
p_trend_egg <- c(
  "0"   = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 0)),
  "10"  = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 10 / 100)),
  "30"  = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 30 / 100)),
  "100" = get_wald_p_pooled(fit_mod4b_full,
                            coef_names = c("as.numeric(egg_freq4)", "meat_gram_ea100:as.numeric(egg_freq4)"),
                            coef_wt    = c(1, 100 / 100))
)

# HRs for egg freq at meat intake levels ----------------------------------

meat_values_gramday <- c(0, 10, 30, 100)
egg_hr_table <- purrr::map_dfr(meat_values_gramday, ~ egg_hr_at_meat(fit_mod4a_full, .x))

egg_hr_wide <- egg_hr_table %>%
  mutate(
    meat_gramday = as.numeric(sub(" g/d", "", Meat)),
    cell = sprintf("%.2f (%.2f, %.2f)", HR, Lower, Upper)
  ) %>%
  select(meat_gramday, Egg, cell) %>%
  pivot_wider(names_from = Egg, values_from = cell) %>%
  mutate(None = "1.00 (Ref)") %>%
  select(meat_gramday, None, `1-3/mo`, `1-4/wk`, `5+/wk`) %>%
  arrange(meat_gramday) %>%
  mutate(`P-trend` = fmt_p(p_trend_egg[as.character(meat_gramday)]))

egg_hr_wide %>%
  kable(
    col.names = c("Meat intake (gram/day)", "None", "1-3 times/month",
                  "1-4 times/week", "5+ times/week", "P-trend"),
    align = "lccccc"
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Egg frequency" = 4, " " = 1))

```


## Assumption checking

### Checking the linearity of dietary variables

* To examine whether the associations between meat and other dietary intakes and log hazard were linear, we fit Cox models using restricted cubic splines with 4 knots for each energy-adjusted dietary variable (grams/day). None of the dietary variables showed a significant nonlinear association with incident CVD, except for fruit intake (p = 0.0290).


```{r cox_dietary_check_nonlinear, echo = FALSE, message = FALSE, warning = FALSE}

library(rms)
library(Hmisc)

# --- Step 1: Fix knot locations across all imputed datasets --------------

vars_rcs <- c("meat_gram_ea", "fish_gram_ea", "alldairy2_gram_ea",
              "totalveg_gram_ea", "fruits_gram_ea", "refgrains_gram_ea",
              "whole_mixed_grains_gram_ea", "nutsseeds_gram_ea", "legumes_gram_ea")

stacked_data <- bind_rows(imputed_data_list)

knot_list <- lapply(vars_rcs, function(v) {
  rcspline.eval(stacked_data[[v]], nk = 4, knots.only = TRUE)
})
names(knot_list) <- vars_rcs

# --- Step 2: Fit cph with FIXED knots on each imputed dataset ------------

# Build the rcs terms programmatically with explicit knot locations
rcs_terms <- sapply(vars_rcs, function(v) {
  k <- paste(round(knot_list[[v]], 4), collapse = ",")
  sprintf("rcs(%s, parms = c(%s))", v, k)
})

rcs_formula_rhs <- paste(rcs_terms, collapse = " + ")

mod3_rcs_fm <- as.formula(paste(
  "Surv(agein, ageout, inc_CVD) ~ bene_sex_F + rti_race3 + marital + educyou2 +",
  "bmicat + exercise + sleephrs2 + smokecat6 + alccat + kcal100 +", rcs_formula_rhs
))

fit_mod3_rcs_list <- lapply(imputed_data_list, function(d) {
  dd <<- datadist(d)              # <<- assigns to global env, not local scope
  options(datadist = "dd")
  cph(mod3_rcs_fm, data = d, method = "efron", x = TRUE, y = TRUE)
})

# --- Step 3: Pool coefficients + covariance via Rubin's rules -------------

pooled_rcs <- pool_coef_vcov(fit_mod3_rcs_list)
qbar <- pooled_rcs$qbar
V    <- pooled_rcs$vcov

# --- Step 4: Multi-df Wald chi-square test for nonlinearity --------------

# Tests H0: all NONLINEAR spline coefficients for a given variable = 0
wald_chisq_pooled <- function(qbar, V, coef_names) {
  L <- matrix(0, nrow = length(coef_names), ncol = length(qbar),
              dimnames = list(coef_names, names(qbar)))
  for (i in seq_along(coef_names)) L[i, coef_names[i]] <- 1
  
  Lq   <- L %*% qbar
  LVLt <- L %*% V %*% t(L)
  chisq <- as.numeric(t(Lq) %*% solve(LVLt) %*% Lq)
  df    <- length(coef_names)
  p     <- pchisq(chisq, df = df, lower.tail = FALSE)
  c(chisq = chisq, df = df, p_nonlinear = p)
}

# Nonlinear terms for a given rcs variable are all basis columns EXCEPT the
# first (linear) one -- with 4 knots, rcs() produces 3 basis columns total:
nonlinear_terms <- function(varname, all_names) {
  grep(paste0("^", varname, "'"), all_names, value = TRUE)
}

# --- Step 5: Run nonlinearity test for each rcs variable ------------------
nonlin_results <- lapply(vars_rcs, function(v) {
  nl_terms <- nonlinear_terms(v, names(qbar))
  res <- wald_chisq_pooled(qbar, V, nl_terms)
  data.frame(variable = v, chisq = res["chisq"], df = res["df"], p_nonlinear = res["p_nonlinear"])
}) %>% bind_rows()

nonlin_results %>% 
  as_tibble() %>% 
  knitr::kable(digits = c(0, 2, 0, 4))

```

### Checking the proportional hazards assumption

* The proportional hazards assumption was assessed by plotting scaled Schoenfeld residuals against time for each covariate. The residuals scattered randomly around zero with no discernible trend over time, supporting the assumption

```{r cox_ph_assumption_check, echo = FALSE, message = FALSE, fig.height = 5, fig.width = 10}

# Checking for PH assumptions ---------------------------------------------

# Using the 1st imputed data
imp1 <- imputed_data_list[[1]]

cox_imp1 <- coxph(
  formula(fit_mod4a_full[[1]]),
  data = imp1,
  model = TRUE
)

# zph p-values are too sensitive when you have a large sample...
zph_result <- cox.zph(cox_imp1)

# Scatter plot of the scaled Schoenfeld residuals 
plot_zph_gg <- function(zph_obj, term, ylim = NULL, point_alpha = 0.3, 
                        point_size = 1, spline_df = 4) {
  time_vals <- zph_obj$x
  resid_vals <- zph_obj$y[, term]
  
  df <- data.frame(time = time_vals, resid = resid_vals)
  
  p <- ggplot(df, aes(x = time, y = resid)) +
    geom_point(alpha = point_alpha, size = point_size, color = "gray30") +
    geom_smooth(
      method = "loess",
      se = TRUE,
      color = "red",
      fill = "red",
      alpha = 0.3,        # transparency of the CI ribbon
      linewidth = 1
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    labs(
      x = "Time",
      y = paste("Beta(t) for", term),
      title = term
    ) +
    theme_minimal()
  
  if (!is.null(ylim)) {
    p <- p + coord_cartesian(ylim = ylim)
  }
  
  return(p)
}

# Build individual panels
p_meat <- plot_zph_gg(zph_result, "meat_gram_ea100", ylim = c(-20, 20))
p_egg  <- plot_zph_gg(zph_result, "egg_freq4")

# Combine side by side with patchwork
p_meat + p_egg

```
